Advancing spatially explicit fisheries management with age-structured species distribution models

### 2025-01-30 paper.tex
paper

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage[noblocks]{authblk}
\renewcommand*{\Authsep}{, }
\renewcommand*{\Authand}{, }
\renewcommand*{\Authands}{, }
\renewcommand\Affilfont{\small}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Advancing spatially explicit fisheries management with age-structured species distribution models},
  pdfauthor={Eric Ward; Nick Tolimieri},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

  \date{}
\begin{document}
\maketitle
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, interior hidden, borderline west={3pt}{0pt}{shadecolor}, sharp corners, breakable, boxrule=0pt, enhanced]}{\end{tcolorbox}}\fi

\hypertarget{abstract}{%
\subsection{\texorpdfstring{\emph{Abstract}}{Abstract}}\label{abstract}}

@@

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

@@

\hypertarget{methods}{%
\subsection{Methods}\label{methods}}

\hypertarget{data}{%
\subsubsection{Data}\label{data}}

@@

\hypertarget{spatial-age-models}{%
\subsubsection{Spatial age models}\label{spatial-age-models}}

@@

\hypertarget{validating-future-predictive-ability}{%
\subsubsection{Validating future predictive
ability}\label{validating-future-predictive-ability}}

@@

\hypertarget{using-predictions-to-quantify-bycatch-risk}{%
\subsubsection{Using predictions to quantify bycatch
risk}\label{using-predictions-to-quantify-bycatch-risk}}

@@

\hypertarget{results}{%
\subsection{Results}\label{results}}

@@

\hypertarget{discussion}{%
\subsection{Discussion}\label{discussion}}

@@

\hypertarget{age-structured-spatial-patterns}{%
\paragraph{\texorpdfstring{\emph{Age-Structured Spatial
Patterns}}{Age-Structured Spatial Patterns}}\label{age-structured-spatial-patterns}}

@@

\hypertarget{predictive-skill-and-cohort-tracking}{%
\paragraph{\texorpdfstring{\emph{Predictive skill and cohort
tracking}}{Predictive skill and cohort tracking}}\label{predictive-skill-and-cohort-tracking}}

@@

\hypertarget{managing-bycatch-risk}{%
\paragraph{\texorpdfstring{\emph{Managing bycatch
risk}}{Managing bycatch risk}}\label{managing-bycatch-risk}}

@@

\hypertarget{need-for-reliable-age-data}{%
\paragraph{Need for reliable age
data}\label{need-for-reliable-age-data}}

@@

\hypertarget{future-directions}{%
\paragraph{Future directions}\label{future-directions}}

@@

\hypertarget{acknowledgments}{%
\subsubsection{Acknowledgments}\label{acknowledgments}}

We thank Owen Liu and XX reviewers for constructive comments on the
manuscript.

@@

\hypertarget{references}{%
\subsection{References}\label{references}}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-agostini2006}{}}%
Agostini, V. N., Francis, R. C., Hollowed, A. B., Pierce, S. D., Wilson,
C., \& Hendrix, A. N. (2006). {The relationship between Pacific hake
({\emph{Merluccius productus}}) distribution and poleward subsurface
flow in the California Current System}. \emph{Canadian Journal of
Fisheries and Aquatic Sciences}, \emph{63}(12), 2648--2659.
\url{https://doi.org/10.1139/f06-139}

\leavevmode\vadjust pre{\hypertarget{ref-Anderson2024}{}}%
Anderson, S. C., Ward, E. J., English, P. A., Barnett, L. A. K., \&
Thorson, J. T. (2024). {sdmTMB: An R package for fast, flexible, and
user-friendly generalized linear mixed effects models with spatial and
spatiotemporal random fields}. \emph{bioRxiv}.
\url{https://doi.org/10.1101/2022.03.24.485545}

\leavevmode\vadjust pre{\hypertarget{ref-brooks2017}{}}%
Brooks, E., M., Kristensen, K., Benthem, A. van, K. J.and Magnusson,
Berg, C. W., Nielsen, A., Skaug, H. J., Mächler, M., \& Bolker, M., B.
(2017). {glmmTMB balances speed and flexibility among packages for
zero-inflated generalized linear mixed modeling}. \emph{The R Journal},
\emph{9}(2), 378. \url{https://doi.org/10.32614/rj-2017-066}

\leavevmode\vadjust pre{\hypertarget{ref-cao2017}{}}%
Cao, J., Thorson, J. T., Richards, R. A., \& Chen, Y. (2017).
{Spatiotemporal index standardization improves the stock assessment of
northern shrimp in the Gulf of Maine}. \emph{Canadian Journal of
Fisheries and Aquatic Sciences}, \emph{74}(11), 1781--1793.
\url{https://doi.org/10.1139/cjfas-2016-0137}

\leavevmode\vadjust pre{\hypertarget{ref-ciannelli2004}{}}%
Ciannelli, L., Chan, K.-S., Bailey, K. M., \& Stenseth, N. Chr. (2004).
{Nonadditive effects of the environment on the survival of a large
marine fish population}. \emph{Ecology}, \emph{85}(12), 3418--3427.
\url{https://doi.org/10.1890/03-0755}

\leavevmode\vadjust pre{\hypertarget{ref-elith2009}{}}%
Elith, J., \& Leathwick, J. R. (2009). {Species distribution models:
Ecological explanation and prediction across space and time}.
\emph{Annual Review of Ecology, Evolution, and Systematics},
\emph{40}(1), 677--697.
\url{https://doi.org/10.1146/annurev.ecolsys.110308.120159}

\leavevmode\vadjust pre{\hypertarget{ref-fredston2021}{}}%
Fredston, A., Pinsky, M., Selden, R. L., Szuwalski, C., Thorson, J. T.,
Gaines, S. D., \& Halpern, B. S. (2021). {Range edges of North American
marine species are tracking temperature over decades}. \emph{Global
Change Biology}, \emph{27}(13), 3145--3156.
\url{https://doi.org/10.1111/gcb.15614}

\leavevmode\vadjust pre{\hypertarget{ref-Grandin2024}{}}%
Grandin, C. J., Johnson, K. F., M., E. A., \& M., B. A. (2024).
\emph{{Status of the Pacific Hake (whiting) stock in U.S. and Canadian
waters in 2024. Prepared by the Joint Technical Committee of the U.S.
and Canada Pacific Hake/Whiting Agreement, National Marine Fisheries
Service and Fisheries and Oceans Canada. 246 p.}}

\leavevmode\vadjust pre{\hypertarget{ref-guisan2005}{}}%
Guisan, A., \& Thuiller, W. (2005). {Predicting species distribution:
offering more than simple habitat models}. \emph{Ecology Letters},
\emph{8}(9), 993--1009.
\url{https://doi.org/10.1111/j.1461-0248.2005.00792.x}

\leavevmode\vadjust pre{\hypertarget{ref-dharma2022}{}}%
Hartig, F. (2022). \emph{{DHARMa: Residual diagnostics for hierarchical
(multi-level / mixed) regression models}}.
\url{https://CRAN.R-project.org/package=DHARMa}

\leavevmode\vadjust pre{\hypertarget{ref-hilborn1992}{}}%
Hilborn, R., \& Walters, C. J. (1992). \emph{{Quantitative fisheries
stock assessment}}. Springer US.
\url{https://doi.org/10.1007/978-1-4615-3598-0}

\leavevmode\vadjust pre{\hypertarget{ref-ices2020}{}}%
ICES. (2020). {Workshop on unavoidable survey effort reduction
(WKUSER)}. \emph{ICES}. \url{https://doi.org/10.17895/ICES.PUB.7453}

\leavevmode\vadjust pre{\hypertarget{ref-johnson2019}{}}%
Johnson, K. F., Thorson, J. T., \& Punt, A. E. (2019). {Investigating
the value of including depth during spatiotemporal index
standardization}. \emph{Fisheries Research}, \emph{216}, 126--137.
\url{https://doi.org/10.1016/j.fishres.2019.04.004}

\leavevmode\vadjust pre{\hypertarget{ref-keller2017}{}}%
Keller, A. A., Wallace, J. R., \& Methot, R. D. (2017). \emph{{The
Northwest Fisheries Science Center's West coast groundfish bottom trawl
survey: History design, and description}}. NOAA, U.S. Department of
Commerce. \url{https://doi.org/10.7289/V5/TM-NWFSC-136}

\leavevmode\vadjust pre{\hypertarget{ref-kristensen2016}{}}%
Kristensen, K., Nielsen, A., Berg, C. W., Skaug, H., \& Bell, B. M.
(2016). {TMB: Automatic differentiation and laplace approximation}.
\emph{Journal of Statistical Software}, \emph{70}(5).
\url{https://doi.org/10.18637/jss.v070.i05}

\leavevmode\vadjust pre{\hypertarget{ref-laman2018}{}}%
Laman, E. A., Rooper, C. N., Turner, K., Rooney, S., Cooper, D. W., \&
Zimmermann, M. (2018). {Using species distribution models to describe
essential fish habitat in Alaska}. \emph{Canadian Journal of Fisheries
and Aquatic Sciences}, \emph{75}(8), 1230--1255.
\url{https://doi.org/10.1139/cjfas-2017-0181}

\leavevmode\vadjust pre{\hypertarget{ref-leising2024cciea}{}}%
Leising, A., M., H., N., T., G., W., \& A., H. (2024).

\leavevmode\vadjust pre{\hypertarget{ref-lewison2004}{}}%
Lewison, R. L., Freeman, S. A., \& Crowder, L. B. (2004). {Quantifying
the effects of fisheries on threatened species: the impact of pelagic
longlines on loggerhead and leatherback sea turtles}. \emph{Ecology
Letters}, \emph{7}(3), 221--231.
\url{https://doi.org/10.1111/j.1461-0248.2004.00573.x}

\leavevmode\vadjust pre{\hypertarget{ref-lindgren2015}{}}%
Lindgren, F., \& Rue, H. (2015). {Bayesian spatial modelling
with{\emph{R}}-{\textbf{INLA}}}. \emph{Journal of Statistical Software},
\emph{63}(19). \url{https://doi.org/10.18637/jss.v063.i19}

\leavevmode\vadjust pre{\hypertarget{ref-lindgren2011}{}}%
Lindgren, F., Rue, H., \& Lindström, J. (2011). {An explicit link
between Gaussian fields and Gaussian Markov Random Fields: The
dtochastic partial fifferential rquation spproach}. \emph{Journal of the
Royal Statistical Society Series B: Statistical Methodology},
\emph{73}(4), 423--498.
\url{https://doi.org/10.1111/j.1467-9868.2011.00777.x}

\leavevmode\vadjust pre{\hypertarget{ref-melo-merino2020}{}}%
Melo-Merino, S. M., Reyes-Bonilla, H., \& Lira-Noriega, A. (2020).
{Ecological niche models and species distribution models in marine
environments: A literature review and spatial analysis of evidence}.
\emph{Ecological Modelling}, \emph{415}, 108837.
\url{https://doi.org/10.1016/j.ecolmodel.2019.108837}

\leavevmode\vadjust pre{\hypertarget{ref-methot2013}{}}%
Methot, R. D., \& Wetzel, C. R. (2013). {Stock synthesis: A biological
and statistical framework for fish stock assessment and fishery
management}. \emph{Fisheries Research}, \emph{142}, 86--99.
\url{https://doi.org/10.1016/j.fishres.2012.10.012}

\leavevmode\vadjust pre{\hypertarget{ref-ono2024}{}}%
Ono, K., Katara, I., Eliasen, S. K., Broms, C., Campbell, A., dos Santos
Schmidt, T. C., Egan, A., Hølleland, S. N., Jacobsen, J. A., Jansen, T.,
Mackinson, S., Mousing, E. A., Nash, R. D. M., Nikolioudakis, N.,
Nnanatu, C., Nøttestad, L., Singh, W., Slotte, A., Wieland, K., \&
Olafsdottir, A. H. (2024). {Effect of environmental drivers on the
spatiotemporal distribution of mackerel at age in the Nordic Seas during
2010{-}20}. \emph{ICES Journal of Marine Science}, \emph{81}(7),
1282--1294. \url{https://doi.org/10.1093/icesjms/fsae087}

\leavevmode\vadjust pre{\hypertarget{ref-ono2016space}{}}%
Ono, K., Shelton, A. O., Ward, E. J., Thorson, J. T., Feist, B. E., \&
Hilborn, R. (2016). {Space-time investigation of the effects of fishing
on fish populations}. \emph{Ecological Applications}, \emph{26}(2),
392--406.

\leavevmode\vadjust pre{\hypertarget{ref-phillips2017}{}}%
Phillips, N. D., Reid, N., Thys, T., Harrod, C., Payne, N. L., Morgan,
C. A., White, H. J., Porter, S., \& Houghton, J. D. R. (2017). {Applying
species distribution modelling to a data poor, pelagic fish complex: the
ocean sunfishes}. \emph{Journal of Biogeography}, \emph{44}(10),
2176--2187. \url{https://doi.org/10.1111/jbi.13033}

\leavevmode\vadjust pre{\hypertarget{ref-rcore2024}{}}%
R Core Team. (2024). \emph{{R: A language and environment for
statistical computing}}. R Foundation for Statistical Computing.
\url{https://www.R-project.org/}

\leavevmode\vadjust pre{\hypertarget{ref-rijnsdorp2018}{}}%
Rijnsdorp, A. D., Bolam, S. G., Garcia, C., Hiddink, J. G., Hintzen, N.
T., Denderen, P. D. van, \& Kooten, T. van. (2018). {Estimating
sensitivity of seabed habitats to disturbance by bottom trawling based
on the longevity of benthic fauna}. \emph{Ecological Applications},
\emph{28}(5), 1302--1312. \url{https://doi.org/10.1002/eap.1731}

>>>>

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage[noblocks]{authblk}
\renewcommand*{\Authsep}{, }
\renewcommand*{\Authand}{, }
\renewcommand*{\Authands}{, }
\renewcommand\Affilfont{\small}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Advancing spatially explicit fisheries management with age-structured species distribution models},
  pdfauthor={Eric Ward; Nick Tolimieri},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

  \date{}
\begin{document}
\maketitle

\subsection{\texorpdfstring{\emph{Abstract}}{Abstract}}\label{abstract}

@@

\subsection{Introduction}\label{introduction}

@@

\subsection{Methods}\label{methods}

\subsubsection{Data}\label{data}

@@

\subsubsection{Spatial age models}\label{spatial-age-models}

@@

\subsubsection{Validating future predictive
ability}\label{validating-future-predictive-ability}

@@

\subsubsection{Using predictions to quantify bycatch
risk}\label{using-predictions-to-quantify-bycatch-risk}

@@

\subsection{Results}\label{results}

@@

\subsection{Discussion}\label{discussion}

@@

\paragraph{\texorpdfstring{\emph{Age-Structured Spatial
Patterns}}{Age-Structured Spatial Patterns}}\label{age-structured-spatial-patterns}

@@

\paragraph{\texorpdfstring{\emph{Predictive skill and cohort
tracking}}{Predictive skill and cohort tracking}}\label{predictive-skill-and-cohort-tracking}

@@

\paragraph{\texorpdfstring{\emph{Managing bycatch
risk}}{Managing bycatch risk}}\label{managing-bycatch-risk}

@@

\paragraph{Need for reliable age data}\label{need-for-reliable-age-data}

@@

\paragraph{Future directions}\label{future-directions}

@@

\subsubsection{Acknowledgments}\label{acknowledgments}

We thank Owen Liu and Chris Jordan for constructive comments on the
manuscript.

@@

\subsection{References}\label{references}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-agostini2006}
Agostini, V. N., Francis, R. C., Hollowed, A. B., Pierce, S. D., Wilson,
C., \& Hendrix, A. N. (2006). {The relationship between Pacific hake
({\emph{Merluccius productus}}) distribution and poleward subsurface
flow in the California Current System}. \emph{Canadian Journal of
Fisheries and Aquatic Sciences}, \emph{63}(12), 2648--2659.
\url{https://doi.org/10.1139/f06-139}

\bibitem[\citeproctext]{ref-Anderson2024}
Anderson, S. C., Ward, E. J., English, P. A., Barnett, L. A. K., \&
Thorson, J. T. (2024). {sdmTMB: An R package for fast, flexible, and
user-friendly generalized linear mixed effects models with spatial and
spatiotemporal random fields}. \emph{bioRxiv}.
\url{https://doi.org/10.1101/2022.03.24.485545}

\bibitem[\citeproctext]{ref-brooks2017}
Brooks, E., M., Kristensen, K., Benthem, A. van, K. J.and Magnusson,
Berg, C. W., Nielsen, A., Skaug, H. J., Mächler, M., \& Bolker, M., B.
(2017). {glmmTMB balances speed and flexibility among packages for
zero-inflated generalized linear mixed modeling}. \emph{The R Journal},
\emph{9}(2), 378. \url{https://doi.org/10.32614/rj-2017-066}

\bibitem[\citeproctext]{ref-cao2017}
Cao, J., Thorson, J. T., Richards, R. A., \& Chen, Y. (2017).
{Spatiotemporal index standardization improves the stock assessment of
northern shrimp in the Gulf of Maine}. \emph{Canadian Journal of
Fisheries and Aquatic Sciences}, \emph{74}(11), 1781--1793.
\url{https://doi.org/10.1139/cjfas-2016-0137}

\bibitem[\citeproctext]{ref-ciannelli2004}
Ciannelli, L., Chan, K.-S., Bailey, K. M., \& Stenseth, N. Chr. (2004).
{Nonadditive effects of the environment on the survival of a large
marine fish population}. \emph{Ecology}, \emph{85}(12), 3418--3427.
\url{https://doi.org/10.1890/03-0755}

\bibitem[\citeproctext]{ref-elith2009}
Elith, J., \& Leathwick, J. R. (2009). {Species distribution models:
Ecological explanation and prediction across space and time}.
\emph{Annual Review of Ecology, Evolution, and Systematics},
\emph{40}(1), 677--697.
\url{https://doi.org/10.1146/annurev.ecolsys.110308.120159}

\bibitem[\citeproctext]{ref-fredston2021}
Fredston, A., Pinsky, M., Selden, R. L., Szuwalski, C., Thorson, J. T.,
Gaines, S. D., \& Halpern, B. S. (2021). {Range edges of North American
marine species are tracking temperature over decades}. \emph{Global
Change Biology}, \emph{27}(13), 3145--3156.
\url{https://doi.org/10.1111/gcb.15614}

\bibitem[\citeproctext]{ref-Grandin2024}
Grandin, C. J., Johnson, K. F., M., E. A., \& M., B. A. (2024).
\emph{{Status of the Pacific Hake (whiting) stock in U.S. and Canadian
waters in 2024. Prepared by the Joint Technical Committee of the U.S.
and Canada Pacific Hake/Whiting Agreement, National Marine Fisheries
Service and Fisheries and Oceans Canada. 246 p.}}

\bibitem[\citeproctext]{ref-guisan2005}
Guisan, A., \& Thuiller, W. (2005). {Predicting species distribution:
offering more than simple habitat models}. \emph{Ecology Letters},
\emph{8}(9), 993--1009.
\url{https://doi.org/10.1111/j.1461-0248.2005.00792.x}

\bibitem[\citeproctext]{ref-dharma2022}
Hartig, F. (2022). \emph{{DHARMa: Residual diagnostics for hierarchical
(multi-level / mixed) regression models}}.
\url{https://CRAN.R-project.org/package=DHARMa}

\bibitem[\citeproctext]{ref-hilborn1992}
Hilborn, R., \& Walters, C. J. (1992). \emph{{Quantitative fisheries
stock assessment}}. Springer US.
\url{https://doi.org/10.1007/978-1-4615-3598-0}

\bibitem[\citeproctext]{ref-ices2020}
ICES. (2020). {Workshop on unavoidable survey effort reduction
(WKUSER)}. \emph{ICES}. \url{https://doi.org/10.17895/ICES.PUB.7453}

\bibitem[\citeproctext]{ref-johnson2019}
Johnson, K. F., Thorson, J. T., \& Punt, A. E. (2019). {Investigating
the value of including depth during spatiotemporal index
standardization}. \emph{Fisheries Research}, \emph{216}, 126--137.
\url{https://doi.org/10.1016/j.fishres.2019.04.004}

\bibitem[\citeproctext]{ref-keller2017}
Keller, A. A., Wallace, J. R., \& Methot, R. D. (2017). \emph{{The
Northwest Fisheries Science Center's West coast groundfish bottom trawl
survey: History design, and description}}. NOAA, U.S. Department of
Commerce. \url{https://doi.org/10.7289/V5/TM-NWFSC-136}

\bibitem[\citeproctext]{ref-kristensen2016}
Kristensen, K., Nielsen, A., Berg, C. W., Skaug, H., \& Bell, B. M.
(2016). {TMB: Automatic differentiation and laplace approximation}.
\emph{Journal of Statistical Software}, \emph{70}(5).
\url{https://doi.org/10.18637/jss.v070.i05}

\bibitem[\citeproctext]{ref-laman2018}
Laman, E. A., Rooper, C. N., Turner, K., Rooney, S., Cooper, D. W., \&
Zimmermann, M. (2018). {Using species distribution models to describe
essential fish habitat in Alaska}. \emph{Canadian Journal of Fisheries
and Aquatic Sciences}, \emph{75}(8), 1230--1255.
\url{https://doi.org/10.1139/cjfas-2017-0181}

\bibitem[\citeproctext]{ref-leising2024cciea}
Leising, A., M., H., N., T., G., W., \& A., H. (2024).

\bibitem[\citeproctext]{ref-lewison2004}
Lewison, R. L., Freeman, S. A., \& Crowder, L. B. (2004). {Quantifying
the effects of fisheries on threatened species: the impact of pelagic
longlines on loggerhead and leatherback sea turtles}. \emph{Ecology
Letters}, \emph{7}(3), 221--231.
\url{https://doi.org/10.1111/j.1461-0248.2004.00573.x}

\bibitem[\citeproctext]{ref-lindgren2015}
Lindgren, F., \& Rue, H. (2015). {Bayesian spatial modelling
with{\emph{R}}-{\textbf{INLA}}}. \emph{Journal of Statistical Software},
\emph{63}(19). \url{https://doi.org/10.18637/jss.v063.i19}

\bibitem[\citeproctext]{ref-lindgren2011}
Lindgren, F., Rue, H., \& Lindström, J. (2011). {An explicit link
between Gaussian fields and Gaussian Markov Random Fields: The
dtochastic partial fifferential rquation spproach}. \emph{Journal of the
Royal Statistical Society Series B: Statistical Methodology},
\emph{73}(4), 423--498.
\url{https://doi.org/10.1111/j.1467-9868.2011.00777.x}

\bibitem[\citeproctext]{ref-melo-merino2020}
Melo-Merino, S. M., Reyes-Bonilla, H., \& Lira-Noriega, A. (2020).
{Ecological niche models and species distribution models in marine
environments: A literature review and spatial analysis of evidence}.
\emph{Ecological Modelling}, \emph{415}, 108837.
\url{https://doi.org/10.1016/j.ecolmodel.2019.108837}

\bibitem[\citeproctext]{ref-methot2013}
Methot, R. D., \& Wetzel, C. R. (2013). {Stock synthesis: A biological
and statistical framework for fish stock assessment and fishery
management}. \emph{Fisheries Research}, \emph{142}, 86--99.
\url{https://doi.org/10.1016/j.fishres.2012.10.012}

\bibitem[\citeproctext]{ref-ono2024}
Ono, K., Katara, I., Eliasen, S. K., Broms, C., Campbell, A., dos Santos
Schmidt, T. C., Egan, A., Hølleland, S. N., Jacobsen, J. A., Jansen, T.,
Mackinson, S., Mousing, E. A., Nash, R. D. M., Nikolioudakis, N.,
Nnanatu, C., Nøttestad, L., Singh, W., Slotte, A., Wieland, K., \&
Olafsdottir, A. H. (2024). {Effect of environmental drivers on the
spatiotemporal distribution of mackerel at age in the Nordic Seas during
2010{-}20}. \emph{ICES Journal of Marine Science}, \emph{81}(7),
1282--1294. \url{https://doi.org/10.1093/icesjms/fsae087}

\bibitem[\citeproctext]{ref-ono2016space}
Ono, K., Shelton, A. O., Ward, E. J., Thorson, J. T., Feist, B. E., \&
Hilborn, R. (2016). {Space-time investigation of the effects of fishing
on fish populations}. \emph{Ecological Applications}, \emph{26}(2),
392--406.

\bibitem[\citeproctext]{ref-phillips2017}
Phillips, N. D., Reid, N., Thys, T., Harrod, C., Payne, N. L., Morgan,
C. A., White, H. J., Porter, S., \& Houghton, J. D. R. (2017). {Applying
species distribution modelling to a data poor, pelagic fish complex: the
ocean sunfishes}. \emph{Journal of Biogeography}, \emph{44}(10),
2176--2187. \url{https://doi.org/10.1111/jbi.13033}

\bibitem[\citeproctext]{ref-rcore2024}
R Core Team. (2024). \emph{{R: A language and environment for
statistical computing}}. R Foundation for Statistical Computing.
\url{https://www.R-project.org/}

\bibitem[\citeproctext]{ref-rijnsdorp2018}
Rijnsdorp, A. D., Bolam, S. G., Garcia, C., Hiddink, J. G., Hintzen, N.
T., Denderen, P. D. van, \& Kooten, T. van. (2018). {Estimating
sensitivity of seabed habitats to disturbance by bottom trawling based
on the longevity of benthic fauna}. \emph{Ecological Applications},
\emph{28}(5), 1302--1312. \url{https://doi.org/10.1002/eap.1731}

\bibitem[\citeproctext]{ref-thorson2015}
Thorson, J. T., Shelton, A. O., Ward, E. J., \& Skaug, H. J. (2015).
{Geostatistical delta-generalized linear mixed models improve precision
for estimated abundance indices for West Coast groundfishes}. \emph{ICES
Journal of Marine Science}, \emph{72}(5), 1297--1310.
\url{https://doi.org/10.1093/icesjms/fsu243}

\bibitem[\citeproctext]{ref-tolimieri2023sea}
Tolimieri, N., \& Haltuch, M. (2023). {Sea-level index of recruitment
variability improves assessment model performance for sablefish
/textit{Anoplopoma fimbria}}. \emph{Canadian Journal of Fisheries and
Aquatic Sciences}, \emph{80}(6), 1006--1016.

\bibitem[\citeproctext]{ref-tolimieri2018}
Tolimieri, N., Haltuch, M. A., Lee, Q., Jacox, M. G., \& Bograd, S. J.
(2018). {Oceanographic drivers of sablefish recruitment in the
California Current}. \emph{Fisheries Oceanography}, \emph{27}(5),
458--474. \url{https://doi.org/10.1111/fog.12266}
